---
title: "SOK-1005-assignment-2"
author: “Marte J. Skadssem Haetta Eikelmann”
format: pdf
echo: true
output: true
---

### Prepare to answer the assignment:

```{r}
#removing all previous data to start on a clean sheet.

rm(list=ls())


#loading necessary packages

library(tidyverse)
library(rvest)   
library(janitor)
library(lubridate)
library(readr)
```

### Task 1

```{r}
#scraping the table from the URL

url <- "https://www.motor.no/aktuelt/motors-store-vintertest-av-rekkevidde-pa-elbiler/217132"

table <- url %>%
  read_html() %>% 
  html_nodes("table") %>% 
  html_table() 
head(table[[1]])
```

```{r}
#naming the first table car_test

car_test <- table[[1]]
```

```{r}
car_test <- car_test %>%
  row_to_names(1, remove_rows_above = FALSE) %>% #set the first row to columnnames
  rename("modell" = "Modell (temp. varierte fra 0° til -10°)","wltp_tall" = "WLTP-tall","avvik" = "Avvik","stopp" = "STOPP") #renaming the columns
```

```{r}
#extracting numbers from each string

car_test <- car_test %>%
  mutate(stopp = parse_number(stopp), wltp_tall = parse_number(wltp_tall), avvik = parse_number(avvik)) %>%
  mutate(avvik = avvik/100)
```

```{r}
#plotting the data

car_test %>%
  ggplot(aes(x=wltp_tall,y=stopp, label=modell)) +
  geom_point(color="blue") +
  geom_abline(intercept = 0, slope = 1, size = 0.3, color="red") + #making a 45 degree line
  geom_text(hjust=0.5, vjust=0.5, check_overlap = TRUE) +
  xlim(300,650) +
  ylim(250,650) +
  labs(title="Sammenheng mellom WLTP-tall og antall faktisk kjørte kilometer", x="WLTP-tall", y="Antall kilometer før stopp") +
  theme_classic()
```

### Task 2

```{r}
lm(stopp ~ wltp_tall, data = car_test)

```

*The lm() function is used to fit a linear model based on the dataframe we put into it. The two values that appeare when I run the code represent the linear model's intercept with the y-axis and the rate of increase. Both values are constants, which means that the values can be multiplied with other values in an expression, but do not change itself.*

```{r}
#plotting the data

car_test %>%
  ggplot(aes(x=wltp_tall,y=stopp, label=modell)) +
  geom_point(color="blue") +
  geom_abline(intercept = 0, slope = 1, size = 0.3, color="red") + #making a 45 degree line
  geom_smooth(method = lm) + #plotting the fitted linear model
  geom_text(hjust=0.5, vjust=0.5, check_overlap = TRUE) +
  xlim(300,650) +
  ylim(250,650) +
  labs(title="Sammenheng mellom WLTP-tall og antall faktisk kjørte kilometer", x="WLTP-tall", y="Antall kilometer før stopp") +
  theme_classic()
```
